\LARGE A double-sampling extension of the German National Forest Inventory for design-based small area regression estimation of timber volume resources on forest district levels \\ \vspace{2ex} \large Large scale application to the federal state of Rhineland-Palatinate \vspace{-1ex}

### 2018-02-08 Draft_Form/tex/Materials_and_Methods(2).tex
update 08_02_2018(2)
update 08_02_2018(2)

\begin{table}[H]
	\begin{center}
	\caption{Classification accuracies of the \textit{treespecies} variable before and after calibration. $n_{ref}$: number of terrestrial reference plots. $n_{class}$: number of classified plots.}
	\vspace{0.2cm}
	\label{tab:classacc}
 {\small %
	\begin{tabular}{l|c|c|c|c|c|c|c} %8cols
		\hlineB{1}
		\multirow{2}{*}{Main plot species} & \multicolumn{2}{c|}{Producer's accuracy[\%]} & \multicolumn{2}{c|}{User's accuracy[\%]} & \multirow{2}{*}{nref} & \multicolumn{2}{c}{nclass} \\ 
		\cline{2-3} \cline{4-5} \cline{7-8} & uncalib & calib & uncalib & calib & & uncalib & calib \\
		\hline \hline	
		 Beech       & 29.56 & 22.31 & 46.11 & 47.02 & 883 & 566 & 419 \\
		 Douglas Fir & 33.04 & 24.78 & 47.50 & 48.72 & 230 & 160 & 117 \\
		 Oak         & 47.75 & 11.07 & 24.34 & 48.48 & 289 & 567 & 66 \\
		 Spruce      & 43.63 & 53.15 & 60.81 & 61.13 & 651 & 467 & 566 \\
		 Scots Pine  & 55.87 & 22.91 & 29.59 & 46.07 & 179 & 338 & 89 \\
		 Mixed       & 66.40 & 84.49 & 63.69 & 64.53 & 3152 & 3286 & 4127 \\
		\hline
		 & \multicolumn{4}{l|}{Overall accuracy: 54.83\% (61.96\%)} & 5384 & \multicolumn{2}{c}{5384} \\
		 \hline \hline
	\end{tabular}
}%
	\end{center}
\end{table}

@@

The model with highest \adjrsq{} and lowest RMSE was achieved with the variables \meanheight{}, \meanheight{}$^2$, \stddev{} and \alsyear{} derived from the ALS data as well as the \treespecies{} variable derived from the classification map as main effect terms. Additionally, interaction terms between \meanheight{} and \alsyear{}, \stddev{} and \alsyear{} , \meanheight{} and \stddev{}, and \meanheight{} and \treespecies{} were included. The model yielded an \adjrsq{} of 0.48 and an RMSE of 140.62 \mha{} (46.7\%) on the plot level (table \ref{tab:modacc_modterms}, full model). While the study of \citet{hill2017a} only provide the model evaluations on the plot level, the two-phase estimators described in section \ref{sec:SAestimators} derive and apply the regression coefficients and the residuals on the aggregated cluster level. We thus re-evaluated the model as used in the estimators also on the cluster level and found the model accuracies to be higher than on the plot level (\adjrsq{} of 0.53 and RMSE of 101.61 \mha{} and 33.6\%). A stratification to the ALS acquisition years substantially improved the model accuracy and thus proved to be an effective means in accounting for the artificially introduced noise in the data set due to ALS quality variations and time-gaps between the ALS and the terrestrial survey. However, the stratification led to highly unbalanced datasets when stratifying according to the \treespecies{} variable. For this reason, a species individual modelling within each \alsyear{} stratum remained infeasible, but might have further improved the model accuracy. An additional evaluation of the model residuals within each ALS acquisition year stratum (table \ref{tab:adj_r2_within}) revealed that the model accuracies substantially varied between the strata (table \ref{tab:adj_r2_within}). \adjrsq{} values above the average  were particularly achieved in ALS acquisition years close to the terrestrial survey date (0.59 even up to 0.66 on the cluster level). This effect was substantially diminished when dropping the \alsyear{} variable from the model term.\par

@@

It should be mentioned that there is a major difference between the external and internal model approach concerning the treatment of outliers or leverage points for model fitting. In order to ensure the zero-mean-residual property of the \psmall{} and \extpsynth{} estimator under the internal model approach (section \ref{sec:psmall} and \ref{sec:extpsynth}), all observations of the terrestrial sample locations $x \in s_2$ used for estimation must also be included in the modelling frame. Excluding an observation from the model fit thus requires deleting it also from the sampling frame, which has to be regarded as an interference with the random sampling process the design-based inference relies on. For this reason, a removal of potential outliers or leverage points is, strictly speaking, only justified if the terrestrial response value or the explanatory variable value turns out to be truly erroneous (e.g. typos or erroneous measurements). If this is not the case, the removal of outliers or influential data points might increase the model fit, but to the cost of possible bias for the estimates (bias-variance trade-off).

We conducted a leverage and outlier detection on the plot level for the full regression model applying the usual criteria of \textit{Leverages} and \textit{Cook's Distance} as amongst others described in \citet{fahrmeir2013}. The Cook's Distance values $D_i$ did not exceed a value of 0.019, and thus were far apart from the often cited critical threshold of $D_i > 0.5$. However, the critical threshold of $2p/n$ for the leverage values (i.e. twice the average of the hat matrix' diagonal entries) was exceeded by 10\% (568) of the (5377) observations.

We conducted an Influence Analysis, i.e. leverage / outlier detection, for the 'final model'. We here considered the usually applied criteria of \textit{Leverages} and \textit{Cook's Distance} as amongst others described in Fahrmeir 2013 \{Regression: models, methods and applications. \textit{Springer Science \& Business Media}. page 164-167\}. The critical threshold of $2p/n$ (i.e. twice the average of the hat matrix' diagonal entries), was exceeded by 10\% of the observations. However, only 3\% of these leverage points were assigned to studentized residuals > 1 or < -1. Leaving these 3\% of points out and recalculating the final model lead to a $R^{2}_{adj}$ of 0.494 compared to 0.485 when including them in the regression. The Cook's Distance values $D_i$ did not exceed a value of 0.019, and thus were far apart from the often cited critical threshold of $D_i > 0.5$. Based on these findings, we decided not to remove observations from the modelling data set. We added a paragraph commenting on this issue of influential data points in section 3.3 'Final Regression Model', page 9 RHS line 690-707.

@@

>>>>

\begin{table}[H]
	\begin{center}
		\caption{Classification accuracies of the \textit{treespecies} variable before and after calibration. $n_{ref}$: number of terrestrial reference plots. $n_{class}$: number of classified plots.}
		\vspace{0.2cm}
		\label{tab:classacc}
		{\small %
			\begin{tabular}{l|c|c|c|c} %8cols
				\hlineB{1}
				Main plot species & Producer's accuracy[\%] & User's accuracy[\%] & nref & nclass \\
				\hline \hline	
				Beech       & 22.31 & 47.02 & 883 & 419 \\
				Douglas Fir & 24.78 & 48.72 & 230 & 117 \\
				Oak         & 11.07 & 48.48 & 289 & 66 \\
				Spruce      & 53.15 & 61.13 & 651 & 566 \\
				Scots Pine  & 22.91 & 46.07 & 179 & 89 \\
				Mixed       & 84.49 & 64.53 & 3152 & 4127 \\
				\hline
				& \multicolumn{2}{l|}{Overall accuracy: 61.96\%} & 5384 & 5384 \\
				\hline \hline
			\end{tabular}
		}%
	\end{center}
\end{table}

%\begin{table}[H]
%	\begin{center}
%	\caption{Classification accuracies of the \textit{treespecies} variable before and after calibration. $n_{ref}$: number of terrestrial reference plots. $n_{class}$: number of classified plots.}
%	\vspace{0.2cm}
%	\label{tab:classacc}
% {\small %
%	\begin{tabular}{l|c c|c c|c|c c} %8cols
%		\hlineB{1}
%		\multirow{2}{*}{Main plot species} & \multicolumn{2}{c|}{Producer's accuracy[\%]} & \multicolumn{2}{c|}{User's accuracy[\%]} & \multirow{2}{*}{nref} & \multicolumn{2}{c}{nclass} \\ 
%		\cline{2-3} \cline{4-5} \cline{7-8} & uncalib & calib & uncalib & calib & & uncalib & calib \\
%		\hline \hline	
%		 Beech       & 29.56 & 22.31 & 46.11 & 47.02 & 883 & 566 & 419 \\
%		 Douglas Fir & 33.04 & 24.78 & 47.50 & 48.72 & 230 & 160 & 117 \\
%		 Oak         & 47.75 & 11.07 & 24.34 & 48.48 & 289 & 567 & 66 \\
%		 Spruce      & 43.63 & 53.15 & 60.81 & 61.13 & 651 & 467 & 566 \\
%		 Scots Pine  & 55.87 & 22.91 & 29.59 & 46.07 & 179 & 338 & 89 \\
%		 Mixed       & 66.40 & 84.49 & 63.69 & 64.53 & 3152 & 3286 & 4127 \\
%		\hline
%		 & \multicolumn{4}{l|}{Overall accuracy: 54.83\% (61.96\%)} & 5384 & \multicolumn{2}{c}{5384} \\
%		 \hline \hline
%	\end{tabular}
%}%
%	\end{center}
%\end{table}

@@

The model with highest \adjrsq{} and lowest RMSE was achieved with the variables \meanheight{}, \meanheight{}$^2$, \stddev{} and \alsyear{} derived from the ALS data as well as the \treespecies{} variable derived from the classification map as main effect terms. Additionally, interaction terms between \meanheight{} and \alsyear{}, \stddev{} and \alsyear{}, \meanheight{} and \stddev{}, and \meanheight{} and \treespecies{} were included. The model yielded an \adjrsq{} of 0.48 and an RMSE of 140.62 \mha{} (46.7\%) on the plot level (table \ref{tab:modacc_modterms}, full model). While the study of \citet{hill2017a} only provide the model evaluations on the plot level, the two-phase estimators described in section \ref{sec:SAestimators} derive and apply the regression coefficients and the residuals on the aggregated cluster level. We thus re-evaluated the model as used in the estimators also on the cluster level and found the model accuracies to be higher than on the plot level (\adjrsq{} of 0.53 and RMSE of 101.61 \mha{} and 33.6\%). A stratification to the ALS acquisition years substantially improved the model accuracy and thus proved to be an effective means in accounting for the artificially introduced noise in the data set due to ALS quality variations and time-gaps between the ALS and the terrestrial survey. However, the stratification led to a highly unbalanced dataset when stratifying according to the \treespecies{} variable. For this reason, a species individual modeling within each \alsyear{} stratum remained infeasible, but might have further improved the model accuracy. An additional evaluation of the model residuals within each ALS acquisition year stratum revealed that the model accuracies substantially varied between the strata (table \ref{tab:adj_r2_within}). Values above the overall \adjrsq{} were particularly achieved in ALS acquisition years close to the terrestrial survey date (0.59 to 0.66 on the cluster level). This effect was substantially diminished when dropping the \alsyear{} variable from the model term.\par

@@

Concerning the treatment of outliers or leverage points for model fitting, it should be noted that there is a major difference between the external and internal model approach. In order to ensure the zero-mean-residual property of the \psmall{} and \extpsynth{} estimator under the internal model approach (section \ref{sec:psmall} and \ref{sec:extpsynth}), all observations of the terrestrial sample $s_2$ used for estimation must also be included in the modeling frame. Thus, excluding an observation from the model fit does also require its deletion from the sampling frame, which has to be regarded as an interference with the random sampling process the design-based inference relies on. For this reason, a removal of potential outliers or leverage points is, strictly speaking, only justified if the terrestrial response value or the explanatory variable value turns out to be truly erroneous (e.g. typos or measurement errors). If this is not the case, the removal of outliers or influential data points might increase the model fit, but to the cost of possible bias for the estimates. With respect to these considerations, we conducted an analysis of influential observations \citep[pp. 160--167]{fahrmeir2013} on the plot level for the full regression model. In order to identify implausible observations in the explanatory data, we calculated the \textit{leverage} values and found that the critical threshold of $2p/n$ (i.e. twice the average of the hat matrix' diagonal entries) was exceeded by 10\% of all observations. Further investigation revealed that several leverage points showed unusually large \meanheight{} values associated to their respective timber volume density values. These implausible height values particularly occured in ALSyear acquisition years differing from the terrestrial survey date and were thus much more likely caused by harvesting activities in the sample plot area than erroneous ALS data. A deletion of these plots from the sampling frame was thus not justified due to the reasons given before. Also the remaining leverage points could not be definitely attributed to erroneous auxiliary data and were kept in the data set. The same was true for 10 observation with exceptionally large local density values that were identified as \textit{outliers}.

@@

%% Notes on leverages:
%
% - 568 plots exceed the critical threshold of 2p/n
% - 1/5 (20\%) 
% - leverages that can indeed be assumed to have a significant influence on the regression coefficients particularly were 
%   a) per se small \meanheight{} values associated to small timber volume densities
%   b) unusually large \meanheight{} values associated to small timber volume values. These implausible height values particularly occured in ALSyear acquisition years differing from the terrestrial survey. These leverage points were thus much more likely caused by harvesting activities in the sample plot area than erroneous ALS data. A deletion of these plots from the sampling / modelling frame was thus not justified due to the reasons given before. However, fitting the model without these plots turned out not to have any influence on the regression coefficients and the overall model accuracy.

% x-values at the boundaries (small and large values) usually have a high leverage. It does not mean that they are invalid / erroneous.

%We conducted an Influence Analysis, i.e. leverage / outlier detection, for the 'final model'. We here considered the usually applied criteria of \textit{Leverages} and \textit{Cook's Distance} as amongst others described in Fahrmeir 2013 \{Regression: models, methods and applications. \textit{Springer Science \& Business Media}. page 164-167\}. The critical threshold of $2p/n$ (i.e. twice the average of the hat matrix' diagonal entries), was exceeded by 10\% of the observations. However, only 3\% of these leverage points were assigned to studentized residuals > 1 or < -1. Leaving these 3\% of points out and recalculating the final model lead to a $R^{2}_{adj}$ of 0.494 compared to 0.485 when including them in the regression. The Cook's Distance values $D_i$ did not exceed a value of 0.019, and thus were far apart from the often cited critical threshold of $D_i > 0.5$. Based on these findings, we decided not to remove observations from the modelling data set. We added a paragraph commenting on this issue of influential data points in section 3.3 'Final Regression Model', page 9 RHS line 690-707.
